Comments on Blair and Sambanis, 2020, \"Forecasting Civil Wars: Theory and Structure in an Age of 'Big Data' and Machine Learning\", *JCR*

### 2020-06-10 paper.Rmd
rewrite smoothing section

The most consequential issue that we found is that all AUC-ROC values reported in B&S Tables 1 and 2 are calculated using smoothed ROC curves, not the original, empirical ROC curves. The data in rare events problems like this one restricts the number of distinct true positive rate values. As we show below, the use of  smoothed ROC curves to calculate AUC has a dramatic impact on the results.

Although smoothing might appear to be a simple option to aid in the visual interpretation of ROC curves, and is presented as such by B&S, it actually involves estimating an underlying population ROC curve, using a parametric binormal model in this case. There is a literature in statistics and the medical field on such ROC estimators and the conditions under which they might be preferable to ...

The option to smooth the ROC curves is presented by B&S as a trivial choice to aid visual interpretation, but it actually delves into a broader area of parametric or non-parametric population ROC curve estimation. A common motivating scenario is in the area of medical testing, where test results might only be binary or ordinal, not continuous, where only a small number of tests with verified accuracy might be available, and where there is a clear larger population one wants to infer to. The particular application here deviates in several ways from the conditions under which the properties of these estimators have been studies: rare events, our data is more of a census than independent samples from a much larger population, and the use of non-deterministic random forest models.

Far from being a trivial choice, the question of whether to use ROC estimates rather than the empirical ROC curves is a major decision that dramatically impacts the results. B&S do not offer a justification for departing from the common practice of using the empirical ROC curves.

>>>>

The most consequential issue that we found is that all AUC-ROC values reported in B&S Tables 1 and 2 are calculated using smoothed ROC curves, not the original, empirical ROC curves. The data in rare events problems like this one restricts the number of true positive rate values and leads to very non-smooth ROC curves; B&S refer to smoothing in the context of their Figure 1 with ROC curves as easing interpretation. But in fact smoothing is used to produce all AUC-ROC values they report. Standard practice has been to use empirical, not smoothed ROC curves, both for visualization and when calculating AUCs.

Aside from the visual impact of ROC smoothing, the underlying motivation and methods involved in smoothing are much broader and involve parametric estimation of population ROC curves [@henley:1988; @henley:2014]. The main application is in the evaluation of medical tests, e.g. diagnostic imaging, with early methods dating back several decades and ongoing development of both parametric and non-parametric smoothing methods [e.g. @zou:etal:1997; @pulit:2016]; it does not seem to be established which methods should be preferable in a given application [@henley:2014]. It is also not clear whether these methods and rationales are valid in applications like our present one, e.g. where our data is closer to a census of the population and with well-known dependencies like spatial and temporal correlation, rather than a random sample that is approximately i.i.d.
